VariantSpark: Population Scale Clustering of Genotype Information

### 2015-11-26 doc/publications/journal2014/paper.tex
update R methods

\subsection*{R Implementation}
For our R implementation, we read in a VCF file using {\sc read.table}, and then convert the resulting table to a matrix ({\sc vcfMatrix} with {\sc as.data.frame.matrix}.
%TODO VariantAnnotation reads in the file quicker but causes issue down the track so overall read.table is faster, right ? - can you add this explanation to the supplementary material.
In the supplementary material we benchmark this faster approach against reading the VCF file in using the `VariantAnnotation' package.
We remove the first 9 columns which leaves us with just the variants. As with our \variantSpark{} pre-processing, we convert the strings that represent each allele to a numeric value.
We then transpose the matrix with {\sc t(vcfMatrix)}, which results in a data-structure where each row represents an individual. We convert the matrix to a {\sc big.matrix} object,
as required by the \kMeans{} algorithm from the `biganalytics' package (\url{https://cran.r-project.org/web/packages/biganalytics/index.html}), and then call {\sc bigkmeans} with
the {\sc big.matrix} object and the required number of clusters as arguments.

>>>>

\subsection*{R Implementation}
For our R implementation, we utilize {\sc read.table} to read in a VCF file as a tab-delimited document. As this also captures the first nine columns that don't contain variants, we subsequently remove those columns with {\sc vcf[1:9] <- list()}.
%TODO VariantAnnotation reads in the file quicker but causes issue down the track so overall read.table is faster, right ? - can you add this explanation to the supplementary material.
In the supplementary material we benchmark a slightly faster (by less than one minute) approach of reading the VCF file in using the `VariantAnnotation' package.

As with our \variantSpark{} pre-processing, we convert the strings that represent each allele to a numeric value. This process is identical for both of our R approaches, and essentially consists
of applying our {\sc Hamming} function to the dataframe with {\sc sapply}.
We then transpose the matrix with {\sc t(vcfMatrix)}, which results in a data-structure where each row represents an individual. We convert the matrix to a {\sc big.matrix} object,
as required by the \kMeans{} algorithm from the `biganalytics' package (\url{https://cran.r-project.org/web/packages/biganalytics/index.html}), and then call {\sc bigkmeans} with
the {\sc big.matrix} object and the required number of clusters as arguments.
